### 2013-02-03 branching-transit.tex
WorkingWiki importation

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\var}{\mathop{\mbox{Var}}}

@@

\item This induces a distribution on $T$, the time that a successful excursion took to cross:
assuming, as is asymptotically correct, that the number of arriving lineages is independent of how long it took to get there,
if $\tau_x$ is the random time of arrival of a successful lineage,
\[
 \P\{\tau_x \in dt\} \propto \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
\]
and so

\item Hence the length of the hitchhiking segment on \emph{one side} of the selected site $L$ has distribution
(using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
\begin{align*}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] \\
 & \propto \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
 &= \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ;
\end{align*}
hence, defining $\alpha = - \log(1-s_d) > 0$,
\begin{align*}
 \P\{ L > \ell \} &= \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
\end{align*}
This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential($x\sqrt{2}$) conditioned to be at least $\sqrt{\alpha}$.
Note that the total length of a hitchhiking segment is the sum of two independent copies of this.

>>>>

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\var}{\mathop{\mbox{Var}}}
\newcommand{\Exp}{\mathop{\mbox{Exp}}}

@@

\item This induces a distribution on $T$, the time that a successful excursion took to cross:
assuming, as is asymptotically correct, that the number of arriving lineages is independent of how long it took to get there,
if $\tau_x$ is the random time of arrival of a successful lineage,
\[
 \P\{\tau_x \in dt\} \propto \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) .
\]

\item The length of the hitchhiking segment on \emph{one side} of the selected site $L$ is exponential with rate depending on $T$;
integrating over $T$ we get
(using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
\begin{align*}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] \\
 & \propto \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
 &= \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ;
\end{align*}
hence, defining $\alpha = - \log(1-s_d) > 0$,
\begin{align*}
 \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] = \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
\end{align*}
This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential with rate $x\sqrt{2}$ conditioned to be at least $\sqrt{\alpha}$.
Note that the total length of a hitchhiking segment is the sum of two independent copies of this.

\item We could rescale with $x$ and get a limiting distribution for $\tau_x$:
\begin{align*}
  \E\left[\exp\left( -\ell \frac{\tau_x}{x} \right)\right]
        &= \exp\left\{ - x\sqrt{2}\left(\sqrt{\frac{\ell}{x}+\alpha} - \sqrt{\alpha}\right) \right\} \\
        &\xrightarrow{x\to\infty} \exp\left\{ - \frac{\ell}{\sqrt{2\alpha}} \right\} ,
\end{align*}
implying that $\tau_x/x \xrightarrow{a.s.} 1/\sqrt{2\alpha}$.

\item Ok, but how about the CLT?  Differentiating, we see that $\E[\tau_x]=x/\sqrt{2\alpha}$ and $\var[\tau_x]=x/2^{3/2}\alpha^{3/2}$,
which suggests looking at
\begin{align*}
\E\left[\exp\left( -\ell \left(\frac{\tau_x-x/\sqrt{2\alpha}}{\sqrt{x}} \right) \right)\right]
    &= \exp\left\{ - x\sqrt{2}\left(\sqrt{\frac{\ell}{\sqrt{x}}+\alpha} - \sqrt{\alpha}\right) + \frac{\ell}{\sqrt{2\alpha x}} \right\} \\
    &\simeq \exp\left\{ - x\sqrt{2}\left( \frac{\ell}{2\sqrt{\alpha x}} - \frac{\ell^2}{8\alpha^{3/2}x}  \right) + \frac{\ell}{\sqrt{2\alpha x}} \right\} \\
    &= \exp\left\{ \frac{\ell^2}{2^{5/2}\alpha^{3/2}} \right\} ,
\end{align*}
since $\sqrt{\alpha+\epsilon}=\sqrt{\alpha} +\epsilon/2\sqrt{\alpha}-\epsilon^2/8\alpha^{3/2} + O(\epsilon^3)$.
The Laplace transform of a $N(0,\sigma^2)$ is $\exp(\ell^2\sigma^2/2)$, so
\[
\frac{\tau_x - x/\sqrt{2\alpha}}{\sqrt{x}} \xrightarrow{d} N(0,2^{-3/2}\alpha^{-3/2}) \qquad \mbox{as}\;x\to\infty.
\]

\item Translating this limit law back to $L$, we have that
\[
  \tau \simeq \frac{x}{2\alpha} + 2\sqrt{2\alpha^3x} B,
\]
with $B \sim N(0,1)$; since $L \sim \Exp(\tau)$,
\[
  \frac{x}{\sqrt{2\alpha}} L \simeq \Exp\left(1+\frac{4\alpha^2}{\sqrt{x}}B\right) \xrightarrow{x\to\infty} \Exp(1),
\]
i.e.\ $L$ is approximately a Exponential with rate $\sqrt{2\alpha}/x$.

\item Note: if we wanted to kill the motion at some negative location (returns to original patch)
we'd use the result for BM, on p.296 of Feller.
